Filling in data gaps on refugee children: Statistical imputation for missing count data

### 2023-01-12 paper.qmd
add origin/asylum plot for % of children

---
title: "Filling in data gaps on refugee children: Statistical imputation for missing count data"
author: Jan Beise, Yukun Pei (UNICEF); Sebastian Steinmueller (UNHCR)
format: pdf
execute:
  echo: false
---

@@

```{r, message=FALSE, warning=FALSE}
t.missing.asyregion <- dem_longMissing %>% 
  group_by(asylum_region, missingAge) %>% 
  summarise(total = sum(total),
            nOrigin = n_distinct(origin_country)) %>% 
  mutate(freq.totalEndYear = total/sum(total),
         freq.origin = nOrigin / sum(nOrigin)) %>%
  ungroup()

```{r, message=FALSE, warning=FALSE, out.width="80%", out.height="160%"}
#| label: fig-missingAsy
#| fig-cap: Missingness by region of asylum
#| warning: false
#| error: false
#| echo: false
#| message: false

p.asyregion.missing <- ggplot(data = t.missing.asyregion %>% 
                                                   filter(!is.na(asylum_region)),
                                                 aes(x = `missingAge`, 
                                                     y = freq.totalEndYear*100,
                                                     fill = `missingAge`)) +
  geom_bar( stat="identity") +
  scale_fill_manual(values=c(unhcrPaletteBlack[1], unhcrPaletteBlue[4])) +
  facet_wrap(~  `asylum_region`, ncol = 1, scales = "fixed") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(size = 13),
    #    plot.title = element_text(hjust = 0.5),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 13)) +
     #   panel.grid.major.x = element_blank()) +
    theme(legend.title = element_blank(), legend.position = "bottom") +
    coord_flip()

@@

p.top4.children.ori.asy <-  ggplot(data = dem_longMissing_geoplots,
       aes(x = origin_country, y = childrenProp, colour = asylum_region)) +
  geom_point(shape = 19, aes(size = total, alpha = 0.3)) +  # Add points to the plot
  scale_color_manual(values = c(unhcrPaletteBlue[3], unhcrPaletteGreen[1], unhcrPaletteRed[1], unhcrPaletteBlack[1])) +
  scale_size(guide = "none") +
  scale_alpha(guide = "none")

@@

@@

@@

@fig-missingOri shows the distribution of refugees by origin regions separately for the two subsets of the global refugee population with (left side) and without (right side) age-disaggregated data. The most common origin regions are Sub-Saharan Africa and the MENA region for refugees with available demographic information.Those without demographic data availability have most commonly fled from countries in the Americas and the Asia-Pacific region. The left part of the graph shows that in the available data, the proportion of children is lower in these two regions. This provides a first indication that refugees with available sex/age-disaggregated data are fundamentally different from those without such data, and that we cannot simply assume the same demographic distribution between these two groups.

>>>>

---
title: "Filling in data gaps on refugee children: Statistical imputation for missing count data"
author: Jan Beise, Yukun Pei (UNICEF); Sebastian Steinmueller (UNHCR)
format: pdf
execute:
  echo: false
  warning: false
  message: false
---

@@

@@

p.top4.children.ori.asy <-  ggplot(data = dem_longMissing_geoplots,
       aes(x = origin_country, y = childrenProp*100, colour = asylum_region)) +
  geom_point(shape = 19, aes(size = total, alpha = 0.3)) +  # Add points to the plot
  scale_color_manual(values = c(unhcrPaletteBlue[3], unhcrPaletteGreen[1], unhcrPaletteRed[1], unhcrPaletteBlack[1])) +
  scale_size(guide = "none") +
  scale_alpha(guide = "none") +
  labs(x = "Origin country\n(Dots proportional to population size of country of asylum)", y="Percentage of children in country of asylum") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 13),
        axis.title.x = element_text(size = 13),
        axis.title.y = element_text(size = 13),
        panel.grid.major.x = element_blank()) +
    theme(legend.title = element_blank(), legend.position = "bottom") +
    annotate("text", x = 3.6, y = 52, label = "Syrian refugees\nin Türkiye", size = 3.5) +
  geom_curve(
    data =   tibble(
    x1 = c(3.6),
    x2 = c(3.1),
    y1 = c(50), 
    y2 = c(45.3)), aes(x = x1, y = y1, xend = x2, yend = y2),
    arrow = arrow(length = unit(0.08, "inch")), size = 0.8,
    color = "gray20", curvature = -0.3)

@@

t.missing.asyregion <- dem_longMissing %>% 
  group_by(asylum_region, missingAge) %>% 
  summarise(total = sum(total),
            nOrigin = n_distinct(origin_country)) %>% 
  mutate(freq.totalEndYear = total/sum(total),
         freq.origin = nOrigin / sum(nOrigin)) %>%
  ungroup()

@@

```{r, message=FALSE, warning=FALSE, out.width="80%", out.height="180%"}
#| label: fig-missingAsy
#| fig-cap: Missingness by region of asylum
#| warning: false
#| error: false
#| echo: false
#| message: false

p.asyregion.missing <- ggplot(data = t.missing.asyregion %>% 
                                                   filter(!is.na(asylum_region)),
                                                 aes(x = `missingAge`, 
                                                     y = freq.totalEndYear*100,
                                                     fill = `missingAge`)) +
  geom_bar( stat="identity", width = 0.8, position = position_dodge(width=0.8)) +
  scale_fill_manual(values=c(unhcrPaletteBlack[1], unhcrPaletteBlue[4])) +
  theme_minimal() +
  ylab("Percentage of refugees") +
  facet_wrap(~  `asylum_region`, ncol = 1, scales = "fixed") +
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(size = 13),
    #    plot.title = element_text(hjust = 0.5),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 13),
    strip.text = element_text(size = 13)) +
     #   panel.grid.major.x = element_blank()) +
    theme(legend.title = element_blank(), legend.position = "right") +
    coord_flip()

@@

@fig-missingOri shows the distribution of refugees by origin regions separately for the two subsets of the global refugee population withand without age-disaggregated data. The most common origin regions are Sub-Africa and Asia for refugees with available demographic information.Those without demographic data availability have most commonly fled from countries in the Americas and Asia. The left part of the graph shows that in the available data, the proportion of children is lower in these two regions. This provides a first indication that refugees with available sex/age-disaggregated data are fundamentally different from those without such data, and that we cannot simply assume the same demographic distribution between these two groups. The varying missingness in the data across regions of origin is indirectly related to the way demographic data is more available for countries of asylum in Africa and Asia: Since 72 per cent of refugees are hosted in neighbouring countries of asylum (Global Trends 2021), it is not suprising to see better data availability not only for refugees in, but also for those from these regions.
